HW 2: Energy Demand and CO2 Emissions

### 2022-09-22 HW1/HW1.Rmd
finish hw2 1.3

---
title: "HW 1"
author: "Hayden Atchley"
date: "9 September 2022"
output: pdf_document
---

@@

```{r}
evshare <- cars %>% 
	transmute(Year = `Model Year`,
						Cars = `Production (000)` * 1000) %>% 
	left_join(evs, by = "Year") %>% 
	mutate(ev_share = EVs / Cars)

@@

Finding a logistic growth model:

```{r}
data <- tibble(year = as.numeric(evshare$Year),
							 share = as.numeric(evshare$ev_share))

formula <- share~(exp(a+b*year)/(1+exp(a+b*year)))
formula <- share~(exp(a)/(1+exp(a)))
formula <- share~SSlogis(year, 1, median(year), 1)
# model <- 
	nls(formula,
			# start = list(a = 1, b = 0.1),
			data = data,
			trace = TRUE)
```

>>>>

---
title: "HW 1: System Tools"
author: "Hayden Atchley"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

```{r, fig.height=2.7}
evshare <- cars %>% 
	transmute(Year = `Model Year`,
						Cars = `Production (000)` * 1000) %>% 
	left_join(evs, by = "Year") %>% 
	mutate(ev_share = EVs / Cars)

@@

In order to solve this model, I first rearranged the growth equation:

\[f(t) = \frac{e^{a+bt}}{1+e^{a+bt}} \implies
f(t) + f(t)(e^{a+bt}) = e^{a+bt} \implies
f(t) = e^{a+bt}(1-f(t)) \]
\[\implies \frac{f(t)}{1-f(t)} = e^{a+bt} \implies
\ln\left(\frac{f(t)}{1-f(t)}\right) = a+bt\]

for all $0 < f(t) < 1$. \medskip

Treating \(\ln\left(\frac{f(t)}{1-f(t)}\right)\) as the new dependent variable and plotting:

```{r, fig.height=2.7}
logevshare <- evshare %>% 
	mutate(logshare = log(ev_share),
				 Year = as.numeric(Year))

fitline <- lm(logshare~Year, logevshare)

logevshare %>% 
	ggplot(aes(x = Year, y = logshare)) +
	geom_point() +
	geom_smooth(method = 'lm') +
	scale_x_continuous(
		breaks = min(logevshare$Year):max(logevshare$Year)) +
	theme(panel.grid.minor.x = element_blank()) +
	ylab("Log of EV Share")
```

Where the best-fit line is given by \(\ln\left(\frac{f(t)}{1-f(t)}\right) = -788.6 + 0.3889t,
\ \therefore\  a=-788.6, b=0.3889\). Putting these values back in the original equation gives:

\[f(t) = \frac{e^{-788.6 + 0.3889t}}{1+e^{-788.6 + 0.3889t}}.\]

And adding this function to the original plot and expanding the time range:

```{r}
a <- fitline$coefficients["(Intercept)"]
b <- fitline$coefficients["Year"]

growthfun <- function(x){
	exp(a+b*x)/(1+exp(a+b*x))
}

evshare %>%
	mutate(Year = as.numeric(Year)) %>% 
	ggplot(aes(x = Year, y = ev_share)) +
	geom_point() +
	geom_function(fun = growthfun) +
	xlim(NA,2050) +
	ylab("Share of EVs Sold vs. All Cars")
```

```{r}
time_at_share <- function(share){
	(log(share/(1-share)) - a) / b
}
```

Solving the equation
\[0.5 = \frac{e^{-788.6 + 0.3889t}}{1+e^{-788.6 + 0.3889t}}\]
gives a value of `r time_at_share(0.5) %>% round(1)`, which is the time at which EV sales account for half of all car sales.

\newpage

# 4
The Annual Energy Outlook's estimation for EV penetration is given as:

![Share of EV sales (AEO estimate).](image/AEO_estimate.png)

This is a much shallower prediction, i.e. EVs do not penetrate nearly as quickly.

### 2022-09-22 HW2/HW2.Rmd
finish hw2 1.3

```{r}
allem <- vmlong %>% 
	filter(category %in% c("ALL LIGHT DUTY VEHICLES/2", "ALL HDV"),
				 ITEM %in% c("Fuel consumed   (thousand gallons)")) %>% 
	pivot_wider(names_from = ITEM) %>% 
	mutate("Emissions (g CO$_2$/gal)" = c(rep(co2gal[2], 2),
																				rep(co2gal[1], 2)),
				 category = ifelse(str_detect(category, "LIGHT"),
				 									"ALL LDV", category)) %>% 
	arrange(desc(category), desc(YEAR)) %>% 
	mutate("Emissions (g CO$_2$)" = 
				 	`Fuel consumed   (thousand gallons)` *
				 	`Emissions (g CO$_2$/gal)` * 1000,
				 "Emissions (M tons CO$_2$)" = `Emissions (g CO$_2$/gal)` / 1000000)

allem %>% 
	kbl(booktabs = TRUE, escape = FALSE,
			caption = "Emissions of LDVs and HDVs") %>% 
	kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

>>>>

```{r}
allem <- vmlong %>% 
	filter(category %in% c("ALL LIGHT DUTY VEHICLES/2", "ALL HDV"),
				 ITEM %in% c("Fuel consumed   (thousand gallons)")) %>% 
	pivot_wider(names_from = ITEM) %>% 
	mutate("Emissions (g CO$_2$/gal)" = c(rep(co2gal[2], 2),
																				rep(co2gal[1], 2)),
				 category = ifelse(str_detect(category, "LIGHT"),
				 									"ALL LDV", category)) %>% 
	arrange(desc(category), desc(YEAR)) %>% 
	mutate("Emissions (g CO$_2$)" = 
				 	`Fuel consumed   (thousand gallons)` *
				 	`Emissions (g CO$_2$/gal)` * 1000,
				 "Emissions (M tons CO$_2$)" = `Emissions (g CO$_2$)` / 1000000)

allem %>% 
	kbl(booktabs = TRUE, escape = FALSE,
			caption = "Emissions of LDVs and HDVs") %>% 
	collapse_rows(columns = 2, valign = "middle") %>% 
	kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

Calculating the percentage of emissions from HDVs per year:

```{r}
pctem <- allem %>% 
	select(YEAR, category, `Emissions (M tons CO$_2$)`) %>% 
	pivot_wider(names_from = category, values_from = `Emissions (M tons CO$_2$)`) %>% 
	transmute(Year = YEAR,
						"LDV Emissions" = `ALL LDV`,
						"HDV Emissions" = `ALL HDV`,
						"HDV Proportion" = `HDV Emissions` /
							(`HDV Emissions` + `LDV Emissions`))

pctem %>% 
	kbl(booktabs = TRUE, escape = FALSE,
			caption = "HDV Emissions Proportion") %>% 
	kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

From this we can see that HDVs made up a greater percentage of CO~2~ emissions in 2020 than in 2019.

## 1.4
